Light versus dark in strong-lens galaxies: Dark matter haloes that are rounder than their stars

### 2015-08-27 shapes/paper.tex
Added captions; some (incomplete) notes in the text what needs to be done; changed one Figure slightly to be in degrees instead of radians

Weak lensing can also be used to probe the shape and alignment between the luminous and dark matter distributions, but only for `stacked' galaxies \citep{2000astro.ph..6281B,2000ApJ...538L.113N}. \citet{2004ApJ...606...67H} applied this idea to data from the Red-Sequence Cluster Survey (RCS) to measure the first weak lensing signal of halo flattening. They found that dark matter haloes appear to be rounder than the stellar mass distribution, with some weak evidence for alignment. Both measurements are challenging however, and more recent data appear to be at odds with this early work, at least for galaxies in general. They however are in agreement concerning `red' galaxies, which are predominantly early-type galaxies, also favouring dark matter haloes that are more elliptical than their stellar mass components \citep{2006MNRAS.370.1008M,2007ApJ...669...21P,2012A&A...545A..71V}.

@@

Lenses are modelled as free-form mass distributions, consisting of mass tiles or pixels. Such lens models are also called non-parametric, which really just means that many more parameters than data constraints are used. Thus the lens models are non-unique, and hence in fact ensembles of these free-form mass distributions are used. From such a mass map, an inertia tensor is defined as
\begin{equation}\label{eq:inertiatensor}
\begin{pmatrix}
 \sum_\theta M(\boldsymbol{\theta})\theta^{2}_{y} & -\sum_\theta M(\boldsymbol{\theta})\theta_{x}\theta_{y} \\
-\sum_\theta M(\boldsymbol{\theta})\theta_{x}\theta_{y} & \sum_\theta M(\boldsymbol{\theta})\theta^{2}_{x}
\end{pmatrix}
\end{equation}
where the sum is over mass pixels, and $M(\boldsymbol{\theta})$ is the mass in a pixel. The eigenvectors of this inertia tensor give the ellipticity axes, and our ellipticity estimator is
\begin{equation}\label{eq:shapeestimate}
    e \equiv 1 - \frac{\lambda_{2}}{\lambda_{1}},
\end{equation}
where $\lambda_{1}$ and $\lambda_{2}$ ($\lambda_{1} \geq \lambda_{2}$) are the eigenvalues. We use this ellipticity estimators to quantify the shapes both of the stellar-mass distribution and of the dark-matter distribution. For the dark matter, only the mass distribution out to the radius of the outermost image is used; for the stellar part, the distribution with within different multiples of $R_e$ (defined as the half-light radius measured with respect to the Petrosian flux) is used. It is necessary to have a cutoff for the regions used to compute the inertia tensor, or else it would be dominated by the outer regions where there is little information. The mass pixels used in this work are at most 4\% of the diameter of the the full mass map. The central region has smaller pixels, to allow for a density cusp at the centre. Each model ensemble had $10^4$ models.

The non-parametric mass models themselves are constructed using the \Glass\ framework for modelling multiple-image lenses \citep{2014MNRAS.445.2181C}. \Glass\ produces ensembles of mass maps, each of which reproduces exactly the observed image positions and image parities; also time delays, if measured. The lensing data must be in the form of point images, which could be quasars or simply point-like features in otherwise extended images, but this was not a problem for the lenses studied in this work. The estimated stellar mass (explained below in Section~\ref{sec:data}) is taken as a lower limit on the total mass. The additional mass that needs to be put on top, in order to reproduce the lensing observables, is interpreted as dark matter. \textbf{(Some of that mass may actually be gas, but very little is expected in the early-type galaxies used in this work.)} The criteria of lensing observables and non-negative dark matter by themselves would leave the mass distribution too under-determined. Hence, an additional prior is used, requiring the mass distribution to be centrally concentrated, with the same centre as the galaxy light (but not necessarily the shape shape). The precise formulation of the prior is given in Section~3 of \cite{2014MNRAS.445.2181C}. Tests of how well the shape estimator can be recovered from lensing data are given in Section~5.2 and Figure~8 of that paper.

@@

\section{Data}\label{sec:data}
Stellar mass maps are taken from \cite{2011ApJ...740...97L}. In that work, the surface brightness distribution of the galaxy images in all available HST bands was first fitted using {\sc Galfit v2.03b} \citep{2002AJ....124..266P}, and with the resulting flux maps stellar mass maps using the synthesis of the stellar-population (SPS) models of \cite{2003MNRAS.344.1000B} were obtained. The population synthesis is marginalised over the star formation epoch and time scales and over the stellar metallicity. The initial mass function (IMF) was taken to be the log-normal form given by \cite{2003PASP..115..763C}; a more bottom-heavy IMF would change the mass normalisation \citep[cf.][]{2014ApJ...793...96S} but not the shape of the inferred stellar mass distribution \citep[unless the IMF presents significant intrinsic deviations locally, see e.g.][]{2015MNRAS.447.1033M}. We note that the effect of recent claims towards a non-universal IMF on stellar mass-to-light ratio (M/L) is still under debate. While for massive elliptical galaxies, dynamical studies seem to point at contributions at the level of the Salpeter IMF or slightly heavier \citep{2013MNRAS.432.1862C}, spectral line strength constraints can still accommodate a wider range of the stellar M/L depending on the functional form of the IMF \citep{2013MNRAS.429L..15F}. For at least one lens in our sample, \textit{Q2237+030}, the lensing estimates on the mass distribution are inconsistent with a Salpeter IMF \citep{2010MNRAS.409L..30F}. We additionally tested varying the IMF for one lens ({\it B1422+231}), finding that Salpeter IMF is too heavy and gives no solutions. We will perform a more systematic study of the IMF and varying stellar M/L in a forthcoming publication.

@@

\section{Results}\label{sec:results}
\begin{figure*}
  \centering
  \includegraphics[width=.8\linewidth]{Figures/kappaplot.pdf}
  \caption[width=\linewidth]{\textbf{CAPTION}}
  \label{fig:kappaplot}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=.75\linewidth]{Figures/wedges_shears.pdf}
  \caption[width=\linewidth]{The reconstructed lenses are displayed according to the reconstructed total shapes of the dark matter halo $s_{dm}$ and the stellar component $s_{*}$. The wedges show the alignment of the semi-major axes of the components. In case of alignment of the semi-major axes (Aligned), the wedges are vertical. In case of alignment of the semi-major with the semi-minor axes (Anti-Aligned), the wedges are horizontal. The opening angle displays the statistical error in the shape estimate of the dark matter halo. Both the errors in the alignment and the ellipticity of the dark matter distribution denote the ranges $68\%$ of all the mass models lie within. The lenses are additionally color-coded in terms of the required external shear in modelling the lens. \textbf{update}}
  \label{fig:wedgesall}
\end{figure*}

\begin{table*}
  \begin{center}
    \begin{tabular}{l | l l l l l l}
      Lens & $s_{*}$ & $s_{dm}$ & $e_{*}$ & $e_{dm}$ & $\frac{e_{dm}}{e_{*}}$ & $\frac{e_{dm,t}}{e_{*}}$ \\ \hline
      0047 & 0.27 & 0.04$\pm$0.02 & 0.15 & 0.02$\pm$0.01 & 0.13$\pm$0.07 & 0.10$\pm$0.07 \\
      0414 & 0.19 & 0.06$\pm$0.03 & 0.10 & 0.03$\pm$0.01 & 0.30$\pm$0.14 & 0.20$\pm$0.11 \\
      0712 & 0.43 & 0.08$\pm$0.03 & 0.28 & 0.04$\pm$0.01 & 0.15$\pm$0.05 & 0.15$\pm$0.05 \\
      0911 & 0.22 & 0.07$\pm$0.02 & 0.12 & 0.04$\pm$0.01 & 0.29$\pm$0.10 & 0.27$\pm$0.10 \\
      0957 & 0.19 & 0.10$\pm$0.03 & 0.10 & 0.05$\pm$0.01 & 0.49$\pm$0.14 & 0.49$\pm$0.14 \\
      1115 & 0.30 & 0.11$\pm$0.02 & 0.18 & 0.06$\pm$0.01 & 0.32$\pm$0.07 & 0.08$\pm$0.04 \\
      1422 & 0.05 & 0.06$\pm$0.03 & 0.03 & 0.03$\pm$0.01 & 1.03$\pm$0.51 & 0.64$\pm$0.41 \\
      1608 & 0.47 & 0.28$\pm$0.01 & 0.31 & 0.16$\pm$0.01 & 0.53$\pm$0.03 & 0.52$\pm$0.03 \\
      2016 & 0.47 & 0.12$\pm$0.02 & 0.30 & 0.06$\pm$0.01 & 0.21$\pm$0.04 & 0.20$\pm$0.04 \\
      2045 & 0.17 & 0.03$\pm$0.02 & 0.09 & 0.02$\pm$0.01 & 0.19$\pm$0.11 & 0.19$\pm$0.11 \\
      2237 & 0.24 & 0.05$\pm$0.03 & 0.13 & 0.03$\pm$0.01 & 0.21$\pm$0.10 & 0.21$\pm$0.10 \\
    \end{tabular}
    \caption[width=\linewidth]{...}
    \label{tab:ellipratios}
  \end{center}
\end{table*}

@@

@@

@@

\begin{itemize}
\item We found that the dark matter haloes are rounder than the light distribution over the range $1R_e < R < 5R_e$. As we average over larger radii, the lenses, except the ones with a large stellar ellipticity, become increasingly elliptical in their stellar distributions. The same trend can furthermore be seen for the dark matter, although the data are also consistent with no change in shape. The dark matter haloes are never more elliptical than $s_{dm} = 0.2$, while their stars can extend to $s_* > 0.3$.

@@

%Relevant references for misalignments being unstable: 
%- http://adsabs.harvard.edu/abs/1978MNRAS.183..779B : 1978MNRAS.183..779B
%- http://adsabs.harvard.edu/abs/1981MNRAS.196..455B : 1981MNRAS.196..455B
%- http://adsabs.harvard.edu/abs/1979ApJ...233..872H : 1979ApJ...233..872H
%- http://adsabs.harvard.edu/abs/1984ApJ...276..101S : 1984ApJ...276..101S
%- http://adsabs.harvard.edu/abs/1987A%26A...182..202R : 1987A&A...182..202R
%- http://adsabs.harvard.edu/abs/1988A%26A...206..269M : 1988A&A...206..269M
%- http://adsabs.harvard.edu/abs/2007ApJ...670.1027A : 2007ApJ...670.1027A

@@

\section{Shear-Shape degeneracy}\label{sec:shearshapedeg}
\begin{figure*}
  \centering
  \includegraphics[width=.9\linewidth]{Figures/theta_scatter.pdf}
  \caption[width=\linewidth]{\textbf{CAPTION}}
  \label{fig:thetascatter}
\end{figure*}

>>>>

\textbf{Weak lensing can also be used to probe the shape and alignment between the luminous and dark matter distributions, but only for `stacked' galaxies \citep{2000astro.ph..6281B,2000ApJ...538L.113N}. \citet{2004ApJ...606...67H} applied this idea to data from the Red-Sequence Cluster Survey (RCS) to measure the first weak lensing signal of halo flattening. They found that dark matter haloes appear to be rounder than the stellar mass distribution, with some weak evidence for alignment. Both measurements are challenging however, and more recent data appear to be at odds with this early work, at least for galaxies in general. They however are in agreement concerning `red' galaxies, which are predominantly early-type galaxies, also favouring dark matter haloes that are more elliptical than their stellar mass components \citep{2006MNRAS.370.1008M,2007ApJ...669...21P,2012A&A...545A..71V}. [CHANGE THIS PARAGRAPH! Split up into two paragraphs to first describe subtleties of the measurement, and then what the constraints are.]}

@@

Lenses are modelled as free-form mass distributions, consisting of mass tiles or pixels. Such lens models are also called non-parametric, which really just means that many more parameters than data constraints are used. Thus the lens models are non-unique, and hence in fact ensembles of these free-form mass distributions are used. From such a mass map, an inertia tensor is defined as
\begin{equation}\label{eq:inertiatensor}
\begin{pmatrix}
 \sum_\theta M(\boldsymbol{\theta})\theta^{2}_{y} & -\sum_\theta M(\boldsymbol{\theta})\theta_{x}\theta_{y} \\
-\sum_\theta M(\boldsymbol{\theta})\theta_{x}\theta_{y} & \sum_\theta M(\boldsymbol{\theta})\theta^{2}_{x}
\end{pmatrix}
\end{equation}
where the sum is over mass pixels, and $M(\boldsymbol{\theta})$ is the mass in a pixel. The eigenvectors of this inertia tensor give the ellipticity axes, and our ellipticity estimator is
\begin{equation}\label{eq:shapeestimate}
    s \equiv 1 - \frac{\lambda_{2}}{\lambda_{1}},
\end{equation}
where $\lambda_{1}$ and $\lambda_{2}$ ($\lambda_{1} \geq \lambda_{2}$) are the eigenvalues. We use this ellipticity estimators to quantify the shapes both of the stellar-mass distribution and of the dark-matter distribution. For the dark matter, only the mass distribution out to the radius of the outermost image is used; for the stellar part, the distribution with within different multiples of $R_e$ (defined as the half-light radius measured with respect to the Petrosian flux) is used. It is necessary to have a cutoff for the regions used to compute the inertia tensor, or else it would be dominated by the outer regions where there is little information. The mass pixels used in this work are at most 4\% of the diameter of the the full mass map. The central region has smaller pixels, to allow for a density cusp at the centre. Each model ensemble had $10^4$ models.

The non-parametric mass models themselves are constructed using the \Glass\ framework for modelling multiple-image lenses \citep{2014MNRAS.445.2181C}. \Glass\ produces ensembles of mass maps, each of which reproduces exactly the observed image positions and image parities; also time delays, if measured. The lensing data must be in the form of point images, which could be quasars or simply point-like features in otherwise extended images, but this was not a problem for the lenses studied in this work. The estimated stellar mass (explained below in Section~\ref{sec:data}) is taken as a lower limit on the total mass. The additional mass that needs to be put on top, in order to reproduce the lensing observables, is interpreted as dark matter. \textbf{(Some of that mass may actually be gas, but very little is expected in the early-type galaxies used in this work.) [IS THIS SUPPOSED TO BE IN THERE?]} The criteria of lensing observables and non-negative dark matter by themselves would leave the mass distribution too under-determined. Hence, an additional prior is used, requiring the mass distribution to be centrally concentrated, with the same centre as the galaxy light (but not necessarily the shape shape). The precise formulation of the prior is given in Section~3 of \cite{2014MNRAS.445.2181C}. Tests of how well the shape estimator can be recovered from lensing data are given in Section~5.2 and Figure~8 of that paper.

@@

\section{Data}\label{sec:data}
\textbf{CLAUDIO: LOOK UP THE REFERENCES ON THE LENSES TO FIND OUT SPECIFIC MISALIGNMENT/ALIGNMENT ANGLES.}

Stellar mass maps are taken from \cite{2011ApJ...740...97L}. In that work, the surface brightness distribution of the galaxy images in all available HST bands was first fitted using {\sc Galfit v2.03b} \citep{2002AJ....124..266P}, and with the resulting flux maps stellar mass maps using the synthesis of the stellar-population (SPS) models of \cite{2003MNRAS.344.1000B} were obtained. The population synthesis is marginalised over the star formation epoch and time scales and over the stellar metallicity. The initial mass function (IMF) was taken to be the log-normal form given by \cite{2003PASP..115..763C}; a more bottom-heavy IMF would change the mass normalisation \citep[cf.][]{2014ApJ...793...96S} but not the shape of the inferred stellar mass distribution \citep[unless the IMF presents significant intrinsic deviations locally, see e.g.][]{2015MNRAS.447.1033M}. We note that the effect of recent claims towards a non-universal IMF on stellar mass-to-light ratio (M/L) is still under debate. While for massive elliptical galaxies, dynamical studies seem to point at contributions at the level of the Salpeter IMF or slightly heavier \citep{2013MNRAS.432.1862C}, spectral line strength constraints can still accommodate a wider range of the stellar M/L depending on the functional form of the IMF \citep{2013MNRAS.429L..15F}. For at least one lens in our sample, \textit{Q2237+030}, the lensing estimates on the mass distribution are inconsistent with a Salpeter IMF \citep{2010MNRAS.409L..30F}. We additionally tested varying the IMF for one lens ({\it B1422+231}), finding that Salpeter IMF is too heavy and gives no solutions. We will perform a more systematic study of the IMF and varying stellar M/L in a forthcoming publication.

@@

\section{Results}\label{sec:results}
\textbf{Just in general: Should we split up this Sections into subsections?}

\begin{figure*}
  \centering
  \includegraphics[width=.8\linewidth]{Figures/kappaplot.pdf}
  \caption[width=\linewidth]{Convergence $\langle\kappa\rangle$ azimuthally averaged within radii $R$. The green curve shows the contribution to the convergence by the dark matter in the strong lens galaxies reconstructed by \GLASS. The vertical lines denote the radial distances of the source images from the centre of the lens. We note that in the case of \textit{0957}, there are two lensed components of the source galaxy. Thus, the green resp. blue vertical lines correspond to one of the respective components. The dashed black lines show the expected profiles for dark matter halos that follow a NFW density profile. These halo masses are calibrated in order for the values of the convergence to agree at one pixel past the last image radius.}
  \label{fig:kappaplot}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=.75\linewidth]{Figures/wedges_shears.pdf}
  \caption[width=\linewidth]{The reconstructed lenses are displayed according to the reconstructed total shapes of the dark matter halo $s_{dm}$ and the stellar component $s_{*}$. The wedges show the alignment of the semi-major axes of the components. In case of alignment of the semi-major axes (Aligned), the wedges are vertical. In case of alignment of the semi-major with the semi-minor axes (Anti-Aligned), the wedges are horizontal. The opening angle displays the statistical error in the shape estimate of the dark matter halo. Both the errors in the alignment and the ellipticity of the dark matter distribution denote the ranges $68\%$ of all the mass models lie within. The lenses are additionally color-coded in terms of the required external shear in modelling the lens. \textbf{UPDATE}}
  \label{fig:wedgesall}
\end{figure*}

\begin{table*}
  \begin{center}
    \begin{tabular}{l | l l l l l l}
      Lens & $s_{*}$ & $s_{dm}$ & $e_{*}$ & $e_{dm}$ & $\frac{e_{dm}}{e_{*}}$ & $\frac{e_{dm,t}}{e_{*}}$ \\ \hline
      0047 & 0.27 & 0.04$\pm$0.02 & 0.15 & 0.02$\pm$0.01 & 0.13$\pm$0.07 & 0.10$\pm$0.07 \\
      0414 & 0.19 & 0.06$\pm$0.03 & 0.10 & 0.03$\pm$0.01 & 0.30$\pm$0.14 & 0.20$\pm$0.11 \\
      0712 & 0.43 & 0.08$\pm$0.03 & 0.28 & 0.04$\pm$0.01 & 0.15$\pm$0.05 & 0.15$\pm$0.05 \\
      0911 & 0.22 & 0.07$\pm$0.02 & 0.12 & 0.04$\pm$0.01 & 0.29$\pm$0.10 & 0.27$\pm$0.10 \\
      0957 & 0.19 & 0.10$\pm$0.03 & 0.10 & 0.05$\pm$0.01 & 0.49$\pm$0.14 & 0.49$\pm$0.14 \\
      1115 & 0.30 & 0.11$\pm$0.02 & 0.18 & 0.06$\pm$0.01 & 0.32$\pm$0.07 & 0.08$\pm$0.04 \\
      1422 & 0.05 & 0.06$\pm$0.03 & 0.03 & 0.03$\pm$0.01 & 1.03$\pm$0.51 & 0.64$\pm$0.41 \\
      1608 & 0.47 & 0.28$\pm$0.01 & 0.31 & 0.16$\pm$0.01 & 0.53$\pm$0.03 & 0.52$\pm$0.03 \\
      2016 & 0.47 & 0.12$\pm$0.02 & 0.30 & 0.06$\pm$0.01 & 0.21$\pm$0.04 & 0.20$\pm$0.04 \\
      2045 & 0.17 & 0.03$\pm$0.02 & 0.09 & 0.02$\pm$0.01 & 0.19$\pm$0.11 & 0.19$\pm$0.11 \\
      2237 & 0.24 & 0.05$\pm$0.03 & 0.13 & 0.03$\pm$0.01 & 0.21$\pm$0.10 & 0.21$\pm$0.10 \\
    \end{tabular}
    \caption[width=\linewidth]{The ellipticity values of the stellar and dark matter components of the reconstructed strong lens galaxies and their corresponding ratios. The ellipticities $s_{*}$ and $s_{dm}$ are defined by Eq.~\ref{eq:shapeestimate}, while $e_{*}$ and $e_{dm}$ are converted using $e = \frac{\lambda_1-\lambda_2}{\lambda_1+\lambda_2}$. The column $\frac{e_{dm}}{e_{*}}$ contains the values of the ratios of the dark matter relative to the stellar distribution. The column $\frac{e_{dm,t}}{e_{*}}$ displays the corresponding ratios of the ellipticity component of the dark matter distribution projected along the stellar distribution and the stellar ellipticity and takes thus potential misalignments into account. The weak lensing estimates are sensitive only to this combination of the ellipticities. \textbf{MAYBE REFINE THIS A BIT}}
    \label{tab:ellipratios}
  \end{center}
\end{table*}

@@

\textbf{Add here some description on the kappa-plot and that it's basically fine except for 2237 (only bulge is probed). Shall we add the fitted halo masses as well?}

@@

\textbf{Comment on the table with the ellipticity ratios and how it compares to the weak lensing estimates.}

@@

\begin{itemize}
\item \textbf{Probably add here something on the profiles}

\item We found that the dark matter haloes are rounder than the light distribution over the range $1R_e < R < 5R_e$. As we average over larger radii, the lenses, except the ones with a large stellar ellipticity, become increasingly elliptical in their stellar distributions. The same trend can furthermore be seen for the dark matter, although the data are also consistent with no change in shape. The dark matter haloes are never more elliptical than $s_{dm} = 0.2$, while their stars can extend to $s_* > 0.3$.

@@

\textbf{ADD THE COMMENTED REFERENCES ON MISALIGNMENTS BEING UNSTABLE IN THE .tex-FILE IN A NEW PARAGRAPH}
%Relevant references for misalignments being unstable: 
%- http://adsabs.harvard.edu/abs/1978MNRAS.183..779B : 1978MNRAS.183..779B
%- http://adsabs.harvard.edu/abs/1981MNRAS.196..455B : 1981MNRAS.196..455B
%- http://adsabs.harvard.edu/abs/1979ApJ...233..872H : 1979ApJ...233..872H
%- http://adsabs.harvard.edu/abs/1984ApJ...276..101S : 1984ApJ...276..101S
%- http://adsabs.harvard.edu/abs/1987A%26A...182..202R : 1987A&A...182..202R
%- http://adsabs.harvard.edu/abs/1988A%26A...206..269M : 1988A&A...206..269M
%- http://adsabs.harvard.edu/abs/2007ApJ...670.1027A : 2007ApJ...670.1027A

@@

\section{Shear-Shape degeneracy [OTHER TITLE?]}\label{sec:shearshapedeg}
\begin{figure*}
  \centering
  \includegraphics[width=.9\linewidth]{Figures/theta_scatter.pdf}
  \caption[width=\linewidth]{The distribution within the ensemble of reconstructed models in the two-dimensional plane defined by the position angle of the dark matter distribution $\theta_{dm}$ and the direction of the required external shear $\theta_{g}$. Each panel shows the histogram of this two-dimensional distribution, which traces the degeneracy of the two parameters, for each reconstructed strong lens galaxy in the sample. \textbf{[CAPTION ISN'T GREAT, MAYBE...]}}
  \label{fig:thetascatter}
\end{figure*}

\textbf{Add some explanatory text on this Figure. PRASENJIT, COULD YOU COMMENT ON THIS DEGENERACY PLEASE (EXPECTATION, WHAT WE SEE, WHY IT'S FINE...)}
